---
title: "Introduction to the Roller Package"
author: "Rocky Lubbers"
date: "11/29/18"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the Roller Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(roller)
```


## Rolling

The `"roller"` package is used for creating a device with a certain number
of sides and corresponding probabilities, rolling the device a set number 
of times, and visualizing the frequencies of the rolls in the series. 


## Creating a `device`

Step 1: create a `"device"` object with the function `device()`:

```{r}
default_device <- device()
default_device
```

By default, `device()` creates a device with two sides (1, 2), 
and probabilties for each side corresponding to 0.5. The default 
constructor in a sense makes a fair coin ready to be 'rolled'.

That being said, if you want to add more sides or different probabilities, 
you can do that by changing the defatuls with arguments `sides` and `prob`:

```{r}
die <- device(sides = 1:6, prob = rep(1/6, 6))
die
```

You can easily created a loaded die like the following:

```{r}
loaded_die <- device(sides = 1:6, prob = c(0.1, 0.1, 0.1, 0.1, 0.1, 0.5))
loaded_die
```



## Rolling a device

Step 2: After you have created your device object ready to be rolled, you will
use the `roll()` function to do so. The `roll()` function takes in arguments
`device` and `times`, in which you put in your created device and roll it the
amount of `times` you want (by default set to 1). The output is of the class 
`rolls`, and contains a list consisting of the rolls, the sides and probabilities 
of the device, and the total rolls (which is the number of `times` inputted). 

When printed, the output shows the class and the vector of rolls:

```{r}
roll10 <- roll(die, times = 10)
roll10
```


Example - roll the loaded die 10 times:

```{r}
loaded_roll10 <- roll(loaded_die, times = 10)
loaded_roll10
```

You can easily check all the attributes of the roll device with the `summary()`
function, easily displayed in a dataframe for convenience:

```{r}
summary(loaded_roll10)
```



## Plot the Frequencies

The function `plot()` can be used on the roll object to create a graphic
representation of the relative frequencies:

```{r}
plot(roll10)
```


