---
title: "README"
output:
  github_document:
    html_preview: false
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

## Overview

"Roller" is an R package that is used for creating devices
with programmable amounts of sides and probabilities, as well as
rolling the devices and visualizing the frequencies of the rolls.
It provides the following functions:

* `device()` creates a device object (of class `"device"`).
* `roll()` to roll a device object, creating a `"rolls"` object.
* `plot()` method for a `"roll"` object to plot frequencies of a series of rolls.
* `summary()` method for a `"roll"` object.


## Motivation

The motivation for this project was to learn how to create and
operate building an entire R package.


## Installation

Install the development version from GitHub via the package `"devtools"`:

```r
# install.packages("devtools") 
# install "cointoss" (without vignettes)
devtools::install_github("rocpoc/roller")
# install "cointoss" (with vignettes)
devtools::install_github("rocpoc/roller", build_vignettes = TRUE)
```


## Usage
```{r}
library(roller)

# default device
default_device = device()
default_device

# default roll the default device
default_roll = roll(default_device)
default_roll


# create fair die device
die = device(sides = 1:6, prob = rep(1/6, 6))
die

# roll the die 5 times or 100 times
die_roll5 = roll(die, 5)
die_roll100 = roll(die, 100)

die_roll5
die_roll100
```

```{r}
# summary of the 5 roll
summary.rolls(die_roll5)

# summary of the 100 roll
summary.rolls(die_roll100)
```

