---
title: "Lab 2: Vectors and other data structures"
author: Rocky Lubbers
date: 9/7/18
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Data Objects

```{r}
# load data in your R session
load('nba2018-salary-points.RData')
length(player)
head(player)
tail(player)
summary(player)
##
length(salary)
head(salary)
tail(salary)
summary(salary)
##
length(experience)
head(experience)
tail(experience)
summary(experience)
```
## Inspecting the data objects
Do all the chosen objects have the same length?  
- All of the chosen objects have the same length.     
Are there missing values, i.e. NA, in any of the objects?  
- In the objects I chose, there are no missisng (N/A) values.  
Find out what is the class of each of the chosen objects.  
- The class of the *player* object is **character**.  
- The class of the *salary* object is **numeric**.  
- The class of the *experience* object is **character**.  
How do you know if any of the loaded objects is a vector?   
- You use the **is.vector()** function.  
How do you know that a given vector is of a certain data type?  
- You use the **mode()** function.  

---  

## Manipulating Vectors: Subsetting
```{r}
four <- head(player, n = 4)
four[1] # gets first player in vector
four[0] # n/a (does not index, rather gets the data type)
four[-1] # omits first item in vector
four[-c(1,2,3)] # omits everything but the last (= to using negative indices on the first 3 itmes)
four[5] # N/A (no item at this index)
four[c(1,2,2,3,3,3)] # repeats the indexed items as many times as it is called in the vector 
```

---  
```{r}
player[c(seq(from=2, to=length(player), by=2))] #all the even elements in player
salary[c(seq(from=1, to=length(salary), by=2))] #all the odd elements in salary
team[c(seq(from=5, to=length(team), by=5))] #all multiples of 5 (e.g. 5, 10, 15, etc) of team
scored[c(seq(from=10, to=length(scored), by=10))] #elements in positions 10, 20, 30, 40, etc of scored
rev(team[c(seq(from=2, to=length(team), by=2))])#all the even elements in team but this time in reverse order
```
---  

## Logical Subsetting
```{r}
player[team == 'GSW' & position == 'C']  # players in position Center, of Warriors (GSW)
player[team == 'GSW' | team == 'LAL'] # players of both GSW (warriors) and LAL (lakers)
player[team == 'LAL' & (position == 'SG'| position == 'PG') ] #players in positions Shooting Guard and Point Guards LAL
player[team == 'GSW' | team == 'LAL' & position == 'SF'] # subset Small Forwards of GSW and LAL
player[which.max(salary)] # name of the player with largest salary
player[which.min(salary)] # name of the player with smallest salary
player[which.max(scored)] # name of the player with largest number of scored points
salary[which.max(scored)] #salary of the player with largest number of points
max(salary[position =='C']) #largest salary of all Centers
team[which.max(scored)] #team of the player with the largest number of scored points
player[which.max(points3)] #name of the player with the largest number of 3-pointers
```
---  

## Subsetting with Character Vectors
```{r}
library(plotly)
warriors_player = player[team=='GSW']
warriors_salary = salary[team=='GSW']
warriors_points = scored[team=='GSW']
names(warriors_salary) = warriors_player

# Raw Value Plot
plot(warriors_points, warriors_salary)
text(warriors_points, warriors_salary, labels = abbreviate(warriors_player))

#Log scored plot
log_scored <- log(warriors_points)
log_salary <- log(warriors_salary)
plot(log_scored, log_salary)
text(log_scored, log_salary, labels = abbreviate(warriors_player))
```
---  

## Recycling
```{r}
salary_millions = salary / 1000000
scored_points = c(points1 * 1 + points2 * 2 + points3 * 3)
scored_points #scored points is the same as scored
```
---  

## Factors
```{r}
position_fac <- factor(position)
position_fac[team == 'GSW'] #positions of Warriors
position_fac[salary > 15000000] #positions of players with salaries > 15 millions
table(position_fac[salary > 15000000]) #frequencies (counts) of positions with salaries > 15 millions
table(team[position == "SG"]) #relative frequencies (proportions) of ‘SG’ (Shooting Guards) in each team
```
---  

## More Plots
```{r}
plot(scored, salary, col=position_fac, cex = 0.5)
```
---  